---
title: "iteration and list columns"
output: github_document
---

```{r library loading, echo = FALSE, message = FALSE}
library(tidyverse)
library(rvest)
```

## Iteration

### Here's some list
```{r}
l = list(
  vec_numeric = 1:4,
  unif_sample = runif(100),
  mat = matrix(1:8, nrow =2 , ncol =4, byrow = TRUE),
  summary = summary(rnorm(1000))
)

l[['mat']]

l[[4]]

```
Make a list that's hopefully a bit more useful.

```{r}
list_norms = 
  list(
    a = rnorm(20, 0, 5),
    b = rnorm(20, 4, 5),
    c = rnorm(20, 0, 10),
    d = rnorm(20, 4, 10)
  )

list_norms[['b']]
```
Let's reuse the funtion we wrote last time.

```{r}
mean_and_sd = function(x) {
  
  if (!is.numeric(x)) {
    stop("Argument x should be numeric")
  } else if (length(x) == 1) {
    stop("Cannot be computed for length 1 vectors")
  }
  
  mean_x = mean(x)
  sd_x = sd(x)

  out_df = 
    tibble(
      mean = mean_x, 
      sd = sd_x
    )
  return (out_df)
}
```

Let's use the function to take mean and sd of all samples.

```{r}
mean_and_sd(list_norms[["a"]])
mean_and_sd(list_norms[["b"]])
mean_and_sd(list_norms[["c"]])
mean_and_sd(list_norms[["d"]])
```


## Use a for loop

Create output list, and run a for loop
```{r}
output = vector("list", length = 4)

for (i in 1:4) {
  
  output[[i]] = mean_and_sd(list_norms[[i]])
  
}

output
```

## Do the same thing

but with  `map` instead

```{r}
#map(list,function)
output = map(list_norms, mean_and_sd)
#equivalent to the for loop above
```

let's do a couple of other things

```{r}
output = map(list_norms, mean_and_sd)
#outputs in individual df

output = map_dfr(list_norms, mean_and_sd)
#simplifies the output into a more reader-friendly table

output = map_dbl(list_norms, IQR)
#simplifies the output into a more reader-friendly list
```

## List columns!!

```{r}
## easier organizational structure
listcol_df = 
  tibble(
    name = c("a", "b", 'c', "d"),
    samp = list_norms
  )

## can pipe easily as it is a dataframe structure
listcol_df %>% 
  filter(name %in% c("a","b"))

listcol_df %>% 
  select(-samp)
```

```{r}
##This is also a list at the same time; can pull out columns of interest using list syntax
listcol_df[["samp"]][["a"]]
```

Compute mean and sd

```{r}
mean_and_sd(listcol_df[["samp"]][["a"]])
mean_and_sd(listcol_df[["samp"]][["b"]])
mean_and_sd(listcol_df[["samp"]][["c"]])
mean_and_sd(listcol_df[["samp"]][["d"]])


map(listcol_df[['samp']], mean_and_sd)


```

ADD a list column!!

```{r}
listcol_df %>% 
  mutate(
    output = map(samp, mean_and_sd),
    iqr = map_dbl(samp, IQR)
  )

listcol_df %>% 
  mutate(
    output = map(samp, mean_and_sd),
    iqr = map_dbl(samp, IQR)
  ) %>% 
  select(-samp) %>% 
  unnest(output)
```

### NSDUH dataset
output a df by extracting df columns 1,4,5, and put everything into a single df

This is a version of our function las time.
```{r}

nsduh_table_format = function (html, table_num){
  out_table = 
    html |> 
    html_table() |> 
    nth(table_num) |>
    slice(-1) |> 
    select(-contains("P Value"))
  
  return(out_table)
}
```


We need to import html and then extract the corret tables. 
```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

nsduh_html = read_html(nsduh_url)

```


```{r}
nsduh_table_format(nsduh_html, table_num = 1)
nsduh_table_format(nsduh_html, table_num = 4)
nsduh_table_format(nsduh_html, table_num = 5)
```
```{r}
nsduh_df = 
  tibble(
    drug = c("marj", "cocaine", "heroin"),
    table_num = c(1,4,5)
  ) %>% 
  mutate(
    table = map(table_num, nsduh_table_format, html = nsduh_html)
  ) %>% 
  unnest(table)

#similar to lambda function to incorporate html within the actual nsduh_table_format function with backslash x
nsduh_df = 
  tibble(
    drug = c("marj", "cocaine", "heroin"),
    table_num = c(1,4,5)
  ) %>% 
  mutate(
    table = map(table_num, \(x) nsduh_table_format(html = nsduh_html, table_num = x))
  ) %>% 
  unnest(table)
                                            
nsduh_df %>% 
  filter(State == "New York")
```
## Another Example!! 

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USW00022534 = "Molokai_HI",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())

```

Create a list column
```{r}
weather_nest = 
  weather_df %>% 
  nest(data = date:tmin)
  
```

```{r}
#only Central Park data from df
weather_nest[['data']][1]
```
Let's try regressing tmax on tmin

```{r}
lm(tmax ~ tmin, data = weather_nest[['data']][[1]])
lm(tmax ~ tmin, data = weather_nest[['data']][[2]])
lm(tmax ~ tmin, data = weather_nest[['data']][[3]])
```
```{r}
weather_nest %>% 
  mutate(
    model_fit = map(data, \(x) lm(tmax~tmin, data = x))
  ) %>% 
  pull(model_fit)
```





